\t QSpread file format

\j This is the format of a spreadsheet file saved with QSpread. The
same format has to be provided the QSpread when loading through import
filters.

\f
         *
         * fixed format
         * (file header)
         *
         sfh_init equ      $00               ; w initialisation code
         sfh.init equ      $4afc
         sfh_id   equ      $02               ; l application id
         sfh.id   equ      'SPRD'
         sfh_vers equ      $06               ; l version of file format
         sfh.vers equ      '1.00'
         sfh_cols equ      $0a               ; w nr of columns (or 0)
         sfh_rows equ      $0c               ; w nr of rows (or 0)
         sfh.len  equ      $0e               ; header length


\j If the spreadsheet size is given as -1/-1, the current layout size
will be used when loading. The is only used when importing though filters,
to free the filter job from the work to find out the size of the imported file.
Any cells which will fall out of the current size will are therefore not laoded.

\f
         *
         * column width format
         sfw_wdth equ      $00      ; column width in characters


\j The file header information is immediatly followed with list of
the column width. There must be an entry for each column in the sheet.
Entries are allowed to have a minimum of 1 and a maximum of 60 chars.
If the first entry is given as -1, all columns will have the configured
default column width.

\f

         *
         * cell format
         sfc_cnr  equ      $00               ; l cell number (col|row)
         sfc_fwrd equ      $04               ; w cell format word
         sfc_form equ      $06               ; s formular standard string

\j The column width information is followed by the cell information.
The format word is defined bewlow. The order at loading time is not
important, at saving time it is always by row.

\f

         *
         * format word (numeric) definition
         fw..dp   equ      0                   bit 0..4 decimal places
         fw..germ equ      4                   german represtentation
         fw..sign equ      5                   always sign number
         fw..brac equ      6                   parenthesis, when negative
         fw..sepr equ      7                   insert seperators
         fw..just equ      10                  clear right just./ set left just.
         fw..strg equ      11                  exclusivly set for string
         fw..msym equ      12                  bit 12..16 monetary symbol

         *
         * reserved bits in string format
         fw..cont equ      0                   set for continuation string
         fw..bold equ      1                   set if bold on output
         fw..ital equ      2                   set if italic on output
         fw..undl equ      3                   set if underlined on output


         *
         * environment codes
         sfe_id   equ      $00               ; w environment code id
         sfe.id   equ      -1
         sfe_code equ      $02               ; w environment code word
         sfe.nfmt equ      $02               ;        global numeric format
         sfe.sfmt equ      $04               ;        global string format
         sfe.fmtn equ      $06               ;        format numbers on/off
         sfe.ordr equ      $08               ;        entry order
         sfe.auto equ      $0a               ;        auto calculation
         sfe.zero equ      $0c               ;        empty when zero
         sfe_val  equ      $04               ; w environment value


\j Cell and environment format can be freely mixed, and the most recent
entry is the one action taken.


\t Printer driver text meta lanuage

\j To allow great fexibility in printer output, QSpread uses a defined
printer command meta language. This language can't be printed direcly,
but has to be filtered though a special job, which translates the defined
meta codes to printer codes.

\f
         *
         * basic printer meta commands
         * (this is the minimal subset to be supported by any driver)
         *
         cmd_meta equ      27
         cmd_init equ      'init'
         cmd_text equ      'text'
         cmd_newl equ      'newl'
         cmd_newp equ      'newp'
         cmd_bold equ      'bold'
         cmd_ital equ      'ital'
         cmd_undl equ      'undl'
